# KNN Algorithm

### Problem Statement:- 

  - Build a KNN model for Zoo dataset to classify the Animal type.
  
### Data Understanding and preparation

```{r}
library(readxl)
zoo <- read.csv("~/desktop/Digi 360/Module 18/zoo.csv")
head(zoo)
```

```{r}
str(zoo)
```

### Normalization

```{r}
#Normalization
normalize <- function(x) {return ((x - min(x)) / (max(x) - min(x))) }
```

```{r}
zoo.n <- as.data.frame(lapply(zoo[,2:17], normalize))
head(zoo.n)
```

### Splitting the Dataset

```{r}
set.seed(123)
dat.d <- sample(1:nrow(zoo.n),size=nrow(zoo.n)*0.7,replace = FALSE) #random selection of 70% data.
 
train <- zoo[dat.d,] # 70% training data
test <- zoo[-dat.d,] # remaining 30% test data
```

```{r}
head(train[,2:18])
```

```{r}
head(test[,2:18])
```

```{r}
#Creating seperate dataframe for 'Creditability' feature which is our target.
train_labels <- zoo[dat.d,18]
test_labels <-zoo[-dat.d,18]
```

```{r}
head(train_labels)
```

```{r}
head(test_labels)
```

```{r}
# Load class package
library(class)

```

```{r}
#Find the number of observation
NROW(train_labels) 
```

So, we have 149 observations in our training data set. The square root of 70 is between 8 and 9, therefore we’ll create two models. One with ‘K’ value as 8 and the other model with a ‘K’ value as 9.

### Building the model for k = 8 and k = 9

```{r}
knn.8 <- knn(train=train[,2:18], test=test[,2:18], cl=train_labels, k=8)
knn.9 <- knn(train=train[,2:18], test=test[,2:18], cl=train_labels, k=9)

```

```{r}
#Calculate the proportion of correct classification for k = 8, 9
ACC.8 <- 100 * sum(test_labels == knn.8)/NROW(test_labels)
ACC.9 <- 100 * sum(test_labels == knn.9)/NROW(test_labels)
```

```{r}
ACC.8
```

```{r}
ACC.9
```

### Optimizing the k value

```{r}
i=1
k.optm=1
for (i in 1:50){
  knn.mod <- knn(train=train[,2:18], test=test[,2:18], cl=train_labels, k=i)
  k.optm[i] <- 100 * sum(test_labels == knn.mod)/NROW(test_labels)
  k=i
  cat(k,'=',k.optm[i],'')}
```

### Accuracy Plot

```{r}
#Accuracy plot
plot(k.optm, type="b", xlab="K- Value",ylab="Accuracy level")
```

### Model building for k =1

```{r}
knn.1 <- knn(train=train[,2:18], test=test[,2:18], cl=train_labels, k=1)
ACC.1 <- 100 * sum(test_labels == knn.1)/NROW(test_labels)

```

```{r}
ACC.1
```

### Conclusion:- 

  - Model has been built for different K values
  - The highest accuarcy we can see at k = 1.